Objective-First Design of Nanophotonic Waveguide Couplers

### 2011-10-26 presentations/ob1_wg_paper/paper.tex
Added materials for church Bible study.
Revised draft of wg coupler paper (OpEx).

\maketitle
\begin{abstract}
We introduce an ``objective-first'' approach to the design of
    nanophotonic waveguide couplers.
Our method is
    computationally fast (20 minutes on a single-core personal computer), 
    requires no trial-and-error,
    does not require guessing a good starting design,
    can be applied to arbitrary waveguide modes,
    and generates devices with high coupling efficiencies (typically 95\%)
    and small footprints (1-4 square vacuum wavelengths).
\end{abstract}
\ocis{230.7370, 130.3990.}

@@

\begin{figure}[htbp]
    \centering
    \includegraphics[width=\textwidth]{quickfig/intro} 
    \caption{Boundary-value formulation of the design objective.
        The values of $H_z^\text{perfect}$, 
            defined along the dashed box surrounding the design area
            (coupling structure), 
            are shown in orange (color online).
        The values of $H_z^\text{perfect}$ along the top and bottom edges
            of the dashed box are set to zero.
        In this schematic, the fundamental and second-order waveguide modes
            have been chosen as the input and output modes respectively.}
    \label{fig:intro}
\end{figure}

@@

\section{Results}
We apply our method to three practically interesting problems 
    that are hard to solve using other available methods: 
    \begin{enumerate}
    \item a coupler between waveguides of different refractive index and width,
    \item a coupler between waveguide modes of different order and symmetry, and
    \item a coupler between waveguides that confine light 
        using different principles 
        (index guided vs. distributed Bragg reflection guided), 
        i.e., between a slab waveguide and a photonic crystal fiber. 
    \end{enumerate}

The corresponding designs as well as the boundary values which make up the
    respective design objectives ($H_z^\text{perfect}$) 
    are shown in Figs.~\ref{fig:fiber}, \ref{fig:mode} and \ref{fig:aircore}.
As initial designs, we uniformly choose $\epsilon = 9$ throughout 
    the whole computational region.
These results demonstrate that our method
    \begin{itemize}
    \item is computationally fast, 
        requiring only 20 minutes on a single-core personal computer;
    \item generates compact devices with footprints of only
        1-4 square vacuum wavelengths;
    \item generates highly efficient devices with  
        typical coupling efficiencies of $95\%$;
    \item can be applied to arbitrary input and output modes,
        as seen from the diverse selection of desired output waveguide modes;
    \item does not require a good initial design.
        The initial design was simply $\epsilon = 9$ everywhere
            (a somewhat arbitrary guess, other values work as well),
            which in the case of Figs.~\ref{fig:mode} and \ref{fig:aircore}
            produces an initially non-functional device 
            (coupling efficiency of $0\%$).

@@

\begin{figure}[htbp]
    \centering
    \includegraphics[width=\textwidth]{quickfig/fiber} 
    \caption{Coupler from a narrow, high-index ($\epsilon=12.25$) waveguide to
        a wide, low-index ($\epsilon=2.25$) waveguide. 
        On the left, the dielectric structure of the coupler 
        and surrounding input and output waveguides are shown.
        The $H_z^\text{perfect}$ boundary values used as 
        the design objective are plotted on the left as well
        (top and bottom boundary values were set to zero).
        On the right, we show the simulation results used
        to compute the performance of the device.
        The coupler converts $96.3\%$ of the input power to the
        designated output mode.
        The device is also extremely compact, 
        convering only $36 \times 66$ grid points,
        where the vacuum wavelength is 42 grid points.
        Computation time was 20 minutes on a personal computer.}
    \label{fig:fiber}
\end{figure}
\begin{figure}[htbp]
    \centering
    \includegraphics[width=\textwidth]{quickfig/mode-conv} 
    \caption{Coupler that converts the fundamental waveguide mode to the
        second-order waveguide mode
        (a movie of the design progress is included as a
        supplementary material).
        This problem is quite difficult since the two modes are of 
        opposite symmetry.
        For example, adiabatic approaches cannot be applied to this case.
        However, our method produces a device 
        (which has the same dimensions and vacuum wavelength as Fig. 1) 
        with a coupling efficiency of $95.5\%$. 
        Computation time was 20 minutes on a personal computer.
        }
    \label{fig:mode}
    \end{figure}
\begin{figure}[htbp]
    \centering
    \includegraphics[width=\textwidth]{quickfig/air-core}
    \caption{Coupler between a dielectric slab waveguide to 
        an air-core waveguide.
        Here, not only are the modes of opposite symmetry,
        but the output waveguide operates on a fundamentally different
        principle (distributed reflection) than the input waveguide 
        (index guided).
        The device still achieves an efficiency of $83.3\%$, demonstrating the
        versatility of our method.
        The vacuum wavelength is 25 grid points, 
        while the device footprint is still $36 \times 66$ grid points.
        Computation time was 20 minutes on a personal computer.
        }
        \label{fig:aircore}
\end{figure}

@@

We then apply an objective-first approach to the design of 
    three practical nanophotonic waveguide couplers which are difficult, 
    at best, to solve with existing methods. 
We show that our method produces
    high-efficiency designs ($\sim 95\%$ efficiency) 
    in small footprints ($\sim 1$ square vacuum wavelength),
    is computationally fast (20 minutes on a single-core personal computer), and
    does not require trial-and-error, or even 
    a good starting design.\\

>>>>

\maketitle
\begin{abstract}
We introduce an ``objective-first'' approach to the design of
    nanophotonic waveguide couplers.
Our method is
    computationally fast (15 minutes on a single-core personal computer), 
    requires no trial-and-error,
    does not require guessing a good starting design,
    can be applied to arbitrary waveguide modes,
    and generates devices with high coupling efficiencies (typically 95\%)
    and small footprints (1-4 square vacuum wavelengths).
\end{abstract}
\ocis{230.7370, 130.3990.}

@@

\begin{figure}[htbp]
    \centering
    \includegraphics[width=\textwidth]{intro} 
    \caption{Boundary-value formulation of the design objective.
        The values of $H_z^\text{perfect}$, 
            defined along the dashed box surrounding the design area
            (coupling structure), 
            are shown in orange (color online).
        The values of $H_z^\text{perfect}$ along the top and bottom edges
            of the dashed box are set to zero.
        In this schematic, the fundamental and second-order waveguide modes
            have been chosen as the input and output modes respectively.}
    \label{fig:intro}
\end{figure}

@@

\section{Results}
We apply our method to five practically interesting problems 
    that are hard to solve using other available methods: 
    \begin{enumerate}
    \item a coupler between waveguides of different refractive index and width,
    \item a coupler between waveguide modes of different order and symmetry,
    \item a coupler between waveguides that confine light 
        using different principles 
        (index guided vs. distributed Bragg reflection guided), 
        i.e., between a slab waveguide and a photonic crystal fiber. 
    \item a coupler from a dielectric to a plasmonic metal-insulator-metal 
        waveguide, and
    \item a coupler from a dielectric waveguide to a (plasmonic) metal wire.
    \end{enumerate}

The corresponding designs as well as the boundary values which make up the
    respective design objectives ($H_z^\text{perfect}$) 
    are shown in Figs.~\ref{fig:fiber}, \ref{fig:mode} and \ref{fig:aircore}.
As initial designs, we uniformly choose $\epsilon = 9$ throughout 
    the whole computational region.
These results demonstrate that our method
    \begin{itemize}
    \item is computationally fast, 
        requiring only 15 minutes on a single-core personal computer;
    \item generates compact devices with footprints of only
        1-4 square vacuum wavelengths;
    \item generates highly efficient devices with  
        typical coupling efficiencies of $95\%$;
    \item can be applied to arbitrary input and output modes,
        as seen from the diverse selection of desired output waveguide modes;
    \item does not require a good initial design.
        The initial design was simply $\epsilon = 9$ everywhere
            (a somewhat arbitrary guess, other values work as well),
            which in the case of Figs.~\ref{fig:mode} and \ref{fig:aircore}
            produces an initially non-functional device 
            (coupling efficiency of $0\%$).

@@

\begin{figure}[htbp]
    \centering
    \includegraphics[width=\textwidth]{1} 
    \caption{Coupler from a narrow, high-index ($\epsilon=12.25$) waveguide to
            a wide, low-index ($\epsilon=2.25$) waveguide. 
        The $H_z^\text{perfect}$ boundary values used as 
            the design objective are plotted in the upper-left quadrant, 
            the generated structure is shown in the upper-right quadrant, and
            the simulated $H_z$ fields of the device are shown 
            in the bottom two plots.
        The computed efficiency of the coupler is high, $98.2\%$, and
            the device is also extremely compact, 
            convering only $36 \times 66$ grid points,
            where the vacuum wavelength is 42 grid points.
        Computation time was 15 minutes on a personal computer.}
    \label{fig:fiber}
\end{figure}
\begin{figure}[htbp]
    \centering
    \includegraphics[width=\textwidth]{2} 
    \caption{Coupler that converts the fundamental waveguide mode to the
            second-order waveguide mode
            (a movie of the design progress is included as a
            supplementary material).
        This problem is quite difficult since the two modes are of 
            opposite symmetry.
        For example, adiabatic approaches cannot be applied to this case.
        However, our method produces a device 
            (which has the same dimensions and vacuum wavelength as 
            Fig.~\ref{fig:fiber})  
            with a coupling efficiency of $97.0\%$. 
        Computation time was 15 minutes on a personal computer.
        }
    \label{fig:mode}
\end{figure}
\begin{figure}[htbp]
    \centering
    \includegraphics[width=\textwidth]{3}
    \caption{Coupler between a dielectric slab waveguide to 
            an air-core waveguide.
        Here, not only are the modes of opposite symmetry,
            but the output waveguide operates on a fundamentally different
            principle (distributed reflection) than the input waveguide 
            (index guided).
        The device still achieves an efficiency of $85.1\%$, demonstrating the
            versatility of our method.
        The vacuum wavelength is 25 grid points, 
            while the device footprint is still $36 \times 66$ grid points.
        Computation time was 15 minutes on a personal computer.
        }
        \label{fig:aircore}
\end{figure}
\begin{figure}[htbp]
    \centering
    \includegraphics[width=\textwidth]{4}
    \caption{Coupler between a dielectric slab waveguide to 
            a plasmonic metal-insulator-metal waveguide.
        The efficiency of the device is $92.9\%$ and 
            has the same wavelength and footprint as the device in
            Fig.~\ref{fig:aircore}.
        Computation time was 15 minutes on a personal computer.
        }
        \label{fig:mim}
\end{figure}
\begin{figure}[htbp]
    \centering
    \includegraphics[width=\textwidth]{5}
    \caption{Coupler between a dielectric slab waveguide to 
            a plasmonic wire waveguide.
        The efficiency of the device is $93.2\%$ and 
            has the same wavelength and footprint as the device in
            Fig.~\ref{fig:aircore}.
        Computation time was 15 minutes on a personal computer.
        }
        \label{fig:wire}
\end{figure}

@@

We then apply an objective-first approach to the design of 
    five practical nanophotonic waveguide couplers which are difficult, 
    at best, to solve with existing methods. 
We show that our method produces
    high-efficiency designs ($\sim 95\%$ efficiency) 
    in small footprints ($\sim 1$ square vacuum wavelength),
    is computationally fast (15 minutes on a single-core personal computer), and
    does not require trial-and-error, or even 
    a good starting design.\\
